---
title: "Movies Recommendation in Flixster"
author: "Robson Teixeira, Nuno Gomes"
date: "17/05/2020"
output:
  html_document:
    df_print: kable
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
geometry:
- top=25mm
- bottom=25mm
- left=25mm
- right=25mm
- heightrounded
highlight-style: pygments
linkcolor: blue
mainfont: Arial
fontsize: 12pt
sansfont: Verdana
documentclass: report
urlcolor: blue
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(!require(tidyverse)) 
  install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(funModeling)) 
  install.packages("funModeling", repos = "http://cran.us.r-project.org")
if(!require(caret)) 
  install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(Hmisc)) 
  install.packages("Hmisc", repos = "http://cran.us.r-project.org")
if(!require(ggplot2)) 
  install.packages("ggplot2", repos = "http://cran.us.r-project.org")

library(tidyverse)
library(funModeling)
library(caret)
library(Hmisc)
library(ggplot2)

options(scipen=999)
```
# Introduction

## Flixster Dataset

The `Flixster` is a social movie site which allow users to share movie ratings, discover new movies and meet others with similar movie taste. The dataset generated from `Flisxter` site consist of 8196077 ratings of 48794 movies by 147612 users. The data are distributed in three files: 

* `profile.txt` : contains informations regarding the users like user id, gender, age, location, for how long it has been a member

* `movie-names.txt` : contains information regarding movies as name and movie id

* `Ratings.time.txt` : contains information on the ratings given by users to movies and on which date

```{r data_load, echo=FALSE, error=FALSE, warning=FALSE}
users <- read.csv("data/profile.txt", na.strings = c("N/A",""), stringsAsFactors = FALSE)

ratings <- read.table("data/Ratings.timed.txt", header = TRUE, 
                         sep = "\t", skipNul = TRUE,
                         col.names = c("userid", "movieid", "rating", "date"))

movies <- read.delim("data/movie-names.txt", na.strings = c("N/A",""), stringsAsFactors = FALSE )
```

A summary of *movies* is given below, togeher with several first rows of dataframe:

```{r movies_summ, warning=FALSE, error=FALSE, echo=FALSE}
glimpse(movies)
df_status(movies)
head(movies)
```

A summary of *users* is given below, togeher with several first rows of dataframe:

```{r users_summ, warning=FALSE, error=FALSE, echo=FALSE}
glimpse(users)
df_status(users)
head(users)
```

The variables _gender_ and _age_ are presented as character and should be changed to factor and integer respectively.

The variables _location_, _memberfor_, _lastlogin_ and _profileview_ are not needed for the context this analysis, so they should be removed.

There are a significant number of _NA's_ in variables _gender_ (67529 or around 6.73% of total) and _age_ (255618 or around 25.49% of total). We adopt the aproach the replacement the empty values into a new value to treatmet of missing values. The empty values into variable _gender_ will be replaced for new value `unknown`. For the empty values into _age_ variable will be set the value `0`.

A summary of *ratings* is given below, togeher with several first rows of dataframe:

```{r ratings_summ, warning=FALSE, error=FALSE, echo=FALSE}
glimpse(ratings)
df_status(ratings)
head(ratings)
```

# Methods and Analysis

## Data Preparation

In this section we prepare the dataset to be used in the analysis. Some steps of cleaning data are applied.

```{r clean_data, warning=FALSE, message=FALSE}
# Remove no relevant variables
users <- select(users, -c(location, memberfor, lastlogin, profileview) )

# Setting 'gender' as factors and 'age' as integer
users <- users %>% mutate(gender = as.factor(gender)) %>% mutate(age = as.integer(age))

# Add column range_age into user data set
users <- users %>% mutate(range_age = cut_width(age, 10, boundary = 0) )

# replace NA values into gender with "unknown"
users <- users %>% mutate(gender = replace(gender, is.na(gender), "unknown"))

# replace NA values into age with "0"
users <- users %>% mutate(age = replace(age, is.na(age), 0))

flixster <- ratings %>% left_join(users, by = "userid") %>% left_join(movies, by = "movieid")

remove(movies, users, ratings)

```

We split the dataset in two parts, the training set called `train_set` and the test set called `test_set` with 70% and 30% of the original dataset respectively.

```{r split_dataset, warning=FALSE, error=FALSE}

# 'test_set' will be 30% of dataset
set.seed(1, sample.kind="Rounding")
test_index <- createDataPartition(y = flixster$rating, times = 1, p = 0.3, list = FALSE)

train_set <- flixster[-test_index,]
temp <- flixster[test_index,]

# Make sure userid and movieid in 'test_set' are also in 'train_set'
test_set <- temp %>% semi_join(train_set, by = "movieid") %>% semi_join(train_set, by = "userid")

# Add rows removed from 'test_set' set back into 'train_set' set
removed <- anti_join(temp, test_set)
train_set <- rbind(train_set, removed)

rm(test_index, temp, flixster, removed)

```

## Data Exploration

Before start building the model, we need to understand the structure of the data, the distribution of variables and the relationship of the predictors. This information will help build a better model.

```{r}
glimpse(train_set)
```

From this initial exploration, we can see that `train_set` consists of 5757622 observations spread across 8 variables. The user information are stored in userid, gender, age, range_age columns; the movie information are stored in movieid and moviename columns; the rating information are stored in rating and date columns.   

Variables   |Type        |Description
------------|------------|------------------------------
userid      |integer     |User unique identifier
movieid     |integer     |Movie unique identifier
rating      |numeric     |rating movie set by user
date        |factor      |Date that user rating movie
gender      |factor      |User gender
age         |integer     |Age of user
range_age   |factor      |rande age of user
moviename   |character   |movie name


